---
title: "Assignment 1 using R Markdown"
# subtitle: "possible subtitle goes here"
author:
  - Mike Noyes^[<michael.noyes@uconn.edu>; Master's student at
    Department of Mathematics, University of Connecticut.]
date: "`r format(Sys.time(), '%d %B %Y')`"
documentclass: article
papersize: letter
fontsize: 11pt
bibliography: template.bib
biblio-style: datalab
keywords: R Markdown, rnorm
# keywords set in YAML header here only go to the properties of the PDF output
# the keywords that appear in PDF output are set in latex/before_body.tex
output:
  bookdown::pdf_document2
  bookdown::html_document2
abstract: |
    In this problem we look at how the approximation of a normal distribution by the Monte Carlo methods compares to the actual normal distribution.
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## some utility functions, see the source code for details
source("utils_template.R")

## specify the packages needed
pkgs <- c("splines2", "DT", "webshot", "leaflet")
need.packages(pkgs)

## external data can be read in by regular functions,
## such as read.table or load

## get output format in case something needs extra effort
outFormat <- knitr::opts_knit$get("rmarkdown.pandoc.to")
## "latex" or "html"

## for latex and html output
isHtml <- identical(outFormat, "html")
isLatex <- identical(outFormat, "latex")
latex <- ifelse(isLatex, '\\LaTeX\\', 'LaTeX')

## specify global chunk options
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
                      out.width = "90%", fig.align = "center")

```


# Introduction {#sec:intro}

To approximate the normal integral using Monte Carlo methods we use the ‘rnorm’ function in R Studio to simulate random normal values. We approximate the probability of values less than “t” by counting all the randomly generated values less than “t” and dividing by the total number of values, “n”.

We look at several different values of t and a few different values of n. One should see that as t increases the probability of a normal random value falling below t should increase. Also, as n increases the approximation using Monte Carlo methods should become closer and closer to the actual value of the normal integral.


# Math Equations {#sec:math}

The major equation involved in this assignment is the approximation of a standard normal distribution by using Monte Carlo methods. The equation for this approximation is shown below


    phi(t) = 1/n * (sum_{i=1}^n I(X_i<=t))



# Figures and Tables{#sec:figure}

The following figure is a boxplot of the bias at all values of t.


<!-- caption of figure may be defined as follows: -->

(ref:cap-bSplines) Bias at all values of t.

```{r bSplines, echo = TRUE, fig.cap = "(ref:cap-bSplines)", fig.width = 8}
mypnorm <- function(t,n) {
  thecdf <- ecdf(rnorm(n))
  thecdf(t)
}

tt <- c(0.0, 0.67, 0.84, 1.28, 1.65, 2.32, 2.58, 3.09, 3.72)
n <- c(100,1000,10000)
thetable <- rbind(t(sapply(n,function(x) mypnorm(tt,x))),pnorm(tt))
dimnames(thetable) <- list(c(n, Inf),tt)
knitr::kable(thetable, digits=4, caption = "Monte Carlo Error Function Approximation")

trials <- replicate(100,sapply(n,function(x) mypnorm(tt,x)))
str(trials)

finaloutput <- data.frame(ecdf=c(trials),t=tt,n=rep(n,each=length(tt)))
boxplot(ecdf ~ t, data=finaloutput, subset=(n==10000))
```


# Code Chunk {#sec:code}

<!-- note that we actually set eval to be FALSE here so that readers are able to
reproduce this template without Python 3. -->

The following code shows how the normal integral is approximated using all values of t and n. This process can be repeated to get multiple trials. The final output is a matrix with the approximated values at t in the first three columns (n=10^2,10^3,10^4 respectively) and the actual value in the final column.

```{r proglang, eval = FALSE, engine = "R", comment = ">>> "}
# n=10^2
trial1 <- c(rnorm(10^2, mean=0, sd=1))

i=1
t1=0
t2=0
t3=0
t4=0
t5=0
t6=0
t7=0
t8=0
t9=0

for(i in 1:100) {
  if(trial1[i]>0) t1 <- t1+1 else t1 <- t1
}

for(i in 1:100) {
  if(trial1[i]>0.67) t2 <- t2+1 else t2 <- t2
}

for(i in 1:100) {
  if(trial1[i]>0.84) t3 <- t3+1 else t3 <- t3
}

for(i in 1:100) {
  if(trial1[i]>1.28) t4 <- t4+1 else t4 <- t4
}

for(i in 1:100) {
  if(trial1[i]>1.65) t5 <- t5+1 else t5 <- t5
}

for(i in 1:100) {
  if(trial1[i]>2.32) t6 <- t6+1 else t6 <- t6
}

for(i in 1:100) {
  if(trial1[i]>2.58) t7 <- t7+1 else t7 <- t7
}

for(i in 1:100) {
  if(trial1[i]>3.09) t8 <- t8+1 else t8 <- t8
}

for(i in 1:100) {
  if(trial1[i]>3.72) t9 <- t9+1 else t9 <- t9
}

t1/100
t2/100
t3/100
t4/100
t5/100
t6/100
t7/100
t8/100
t9/100

x <- matrix(c(t1/100,t2/100,t3/100,t4/100,t5/100,t6/100,t7/100,t8/100,t9/100),ncol=1)
x

# n=10^3
trial2 <- c(rnorm(10^3, mean=0, sd=1))

i=1
tt1=0
tt2=0
tt3=0
tt4=0
tt5=0
tt6=0
tt7=0
tt8=0
tt9=0

for(i in 1:1000) {
  if(trial2[i]>0) tt1 <- tt1+1 else tt1 <- tt1
}

for(i in 1:1000) {
  if(trial2[i]>0.67) tt2 <- tt2+1 else tt2 <- tt2
}

for(i in 1:1000) {
  if(trial2[i]>0.84) tt3 <- tt3+1 else tt3 <- tt3
}

for(i in 1:1000) {
  if(trial2[i]>1.28) tt4 <- tt4+1 else tt4 <- tt4
}

for(i in 1:1000) {
  if(trial2[i]>1.65) tt5 <- tt5+1 else tt5 <- tt5
}

for(i in 1:1000) {
  if(trial2[i]>2.32) tt6 <- tt6+1 else tt6 <- tt6
}

for(i in 1:1000) {
  if(trial2[i]>2.58) tt7 <- tt7+1 else tt7 <- tt7
}

for(i in 1:1000) {
  if(trial2[i]>3.09) tt8 <- tt8+1 else tt8 <- tt8
}

for(i in 1:1000) {
  if(trial2[i]>3.72) tt9 <- tt9+1 else tt9 <- tt9
}

tt1/1000
tt2/1000
tt3/1000
tt4/1000
tt5/1000
tt6/1000
tt7/1000
tt8/1000
tt9/1000

y <- matrix(c(tt1/1000,tt2/1000,tt3/1000,tt4/1000,tt5/1000,tt6/1000,tt7/1000,tt8/1000,tt9/1000),ncol=1)
y

# n=10^4
trial3 <- c(rnorm(10^4, mean=0, sd=1))

i=1
ttt1=0
ttt2=0
ttt3=0
ttt4=0
ttt5=0
ttt6=0
ttt7=0
ttt8=0
ttt9=0

for(i in 1:10000) {
  if(trial3[i]>0) ttt1 <- ttt1+1 else ttt1 <- ttt1
}

for(i in 1:10000) {
  if(trial3[i]>0.67) ttt2 <- ttt2+1 else ttt2 <- ttt2
}

for(i in 1:10000) {
  if(trial3[i]>0.84) ttt3 <- ttt3+1 else ttt3 <- ttt3
}

for(i in 1:10000) {
  if(trial3[i]>1.28) ttt4 <- ttt4+1 else ttt4 <- ttt4
}

for(i in 1:10000) {
  if(trial3[i]>1.65) ttt5 <- ttt5+1 else ttt5 <- ttt5
}

for(i in 1:10000) {
  if(trial3[i]>2.32) ttt6 <- ttt6+1 else ttt6 <- ttt6
}

for(i in 1:10000) {
  if(trial3[i]>2.58) ttt7 <- ttt7+1 else ttt7 <- ttt7
}

for(i in 1:10000) {
  if(trial3[i]>3.09) ttt8 <- ttt8+1 else ttt8 <- ttt8
}

for(i in 1:10000) {
  if(trial3[i]>3.72) ttt9 <- ttt9+1 else ttt9 <- ttt9
}

ttt1/10000
ttt2/10000
ttt3/10000
ttt4/10000
ttt5/10000
ttt6/10000
ttt7/10000
ttt8/10000
ttt9/10000

z <- matrix(c(ttt1/10000,ttt2/10000,ttt3/10000,ttt4/10000,ttt5/10000,ttt6/10000,ttt7/10000,ttt8/10000,ttt9/10000),ncol=1)
z

# True value at t
w <- c(0.5,0.7486,0.7995,0.8997,0.9505,0.9898,0.9951,0.9990,0.9999)
truevalues <- 1-w

attempt1 <- matrix(c(x,y,z,truevalues),nrow=9,ncol=4)
1-attempt1
```

<!-- The expected result is given as follows -->

```R
>>>   [,1]  [,2]   [,3]   [,4]
 [1,] 0.49 0.509 0.4976 0.5000
 [2,] 0.71 0.760 0.7524 0.7486
 [3,] 0.80 0.801 0.8005 0.7995
 [4,] 0.88 0.907 0.9025 0.8997
 [5,] 0.95 0.939 0.9480 0.9505
 [6,] 0.97 0.989 0.9892 0.9898
 [7,] 0.98 0.996 0.9947 0.9951
 [8,] 0.98 0.999 0.9989 0.9990
 [9,] 1.00 0.999 0.9999 0.9999
```


# Summary {#sec:summary}

In summary, we can see that the Monte Carlo methods do a sufficient job of approximating a standard normal integral. As n increases, the approximation gets closer and closer to the actual values of the integral. This Monte Carlo approximation is a great way of estimation an integral if you don't have access to a z-table.
I am fairly knew to R, so my methods aren't the most efficient. Class today, February 1st, really helped me complete the "Figures and Tables" section of this assignment. All in all, this report helped me learn much more about how to use R Studio.